### 2022-06-27 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_ncbitaxonomy_softwarenote@cb3b8649606ace1b96dd62265adf5afc9e3ad0c9 🚀

In this manuscript, we describe \texttt{NCBITaxonomy.jl}, a Julia
package that provides advanced name matching and error handling
capacities for the reconciliation of taxonomic names to the NCBI
database. This package was used to facilitate the development of the
\emph{CLOVER} (Gibb et al. 2021) database of host-virus associations, by
reconciling the names of viruses and mammals from four different
sources, where all of the issues described above were present. More
recently, it has become part of the automated curation of data for the
\emph{VIRION} (Carlson et al. 2022) database, which automatically
curates an up-to-date, authoritative host-mammal network from dozens of
heterogeneous sources. We describe the core capacities of this package,
and highlight how it enables safe, high-performance name reconciliation.

@@

@@

Name finding is primarily done through the \texttt{taxon} function,
which admits either a unique NCBI identifier (\emph{e.g.}
\texttt{taxon(36219)} for the bogue \emph{Boops boops}), a string
(\texttt{taxon("Boops\ boops")}), or a data frame with a restricted list
of names in order to create a name finder function (see the next
section). The \texttt{taxon} method has additional arguments to perform
fuzzy matching in order to catch possible typos
(\texttt{taxon("Boops\ bops";\ strict=false)}), to perform a lowercase
search (useful when alphanumeric codes are part of the taxon name, like
for some viruses), and to restrict the the search to a specific
taxonomic rank. The \texttt{taxon} function also accepts a
\texttt{preferscientificname} keyword, to prevent matching vernacular
names; the use of this keyword ought to be informed by knowledge about
how the data were entered.

@@

>>>>

In this manuscript, we describe \texttt{NCBITaxonomy.jl}, a Julia
package that provides advanced name matching and error handling
capacities for the reconciliation of taxonomic names to the NCBI
database. This package was used to facilitate the development of the
\emph{CLOVER} (Gibb et al. 2021) database of host-virus associations, by
reconciling the names of viruses and mammals from four different
sources, where all of the issues described above were present. More
recently, it has become part of the automated curation of data for the
\emph{VIRION} (Carlson et al. 2022) database, which automatically
curates an up-to-date, authoritative virome network from dozens of
heterogeneous sources. We describe the core capacities of this package,
and highlight how it enables safe, high-performance name reconciliation.

@@

This software note describes version \texttt{v0.2.3} of the package (we
follow semantic versioning), which works on Julia 1.5 upwards. The
dependencies are all resolved by the package manager at installation,
and (on the user-facing side) include the \texttt{StringDistances.jl}
package, allowing users to experiment with different string matching
methods. As is best practices for Julia packages, a
\texttt{Project.toml} file specifying compatible dependencies versions
is distributed with the package. The code is covered by unit-tests (with
about 98\% coverage), as well as integration tests as part of the
documentation (specifically, a use-case detailing how to clean data from
a biodiversity survey, and a use-case aiming to reconstruct a taxonomic
tree for the Lemuriformes).

@@

Name finding, \emph{i.e.} the matching of an arbitrary string to a
taxonomic identifier, is primarily done through the \texttt{taxon}
function, which admits either a unique NCBI identifier (\emph{e.g.}
\texttt{taxon(36219)} for the bogue \emph{Boops boops}), a string
(\texttt{taxon("Boops\ boops")}), or a data frame with a restricted list
of names in order to create a name finder function (see the next
section). The \texttt{taxon} method has additional arguments to perform
fuzzy matching in order to catch possible typos
(\texttt{taxon("Boops\ bops";\ strict=false)}), to perform a lowercase
search (useful when alphanumeric codes are part of the taxon name, like
for some viruses), and to restrict the the search to a specific
taxonomic rank. The \texttt{taxon} function also accepts a
\texttt{preferscientificname} keyword, to prevent matching vernacular
names; the use of this keyword ought to be informed by knowledge about
how the data were entered.

@@

\hypertarget{conclusion}{%
\section{Conclusion}\label{conclusion}}

\texttt{NCBITaxonomy.jl} enables rapid, taxonomically-restricted,
adaptive matching for taxonomic names. By implementing various
combinations of search strategies, it allows users to (i) optimize the
speed of their queries and (ii) avoid usual caveats of simple string
matching. Through explicit exceptions, it allows to write code that will
handle the possible edge cases that cannot be solved automatically in a
way that does not interrupt execution, or requires manual input by the
user. Given the breadth of the NCBI taxonomy database,
\texttt{NCBITaxonomy.jl} is particularly suited to the name cleaning of
large datasets of names.
